Conflict-Free Vertex Coloring Of Planar Graphs

### 2017-02-27 paper/paper.tex
Adding more background, section draft almost done

A graph, denoted $G=(V,E)$, is an ordered pair of two sets: a set of vertices $V$ and a set of edges $E$. Each edge consists of a pair of vertices from $V$. For example, $(u,w) \in E$ is an edge connecting vertices $u$ and $w$ where $u,w \in V$. Vertices are adjacent if they are connected by an edge. A graph can be non-technically described as a set of points (vertices) with lines (edges) connecting them. A simple graph is an undirected graph where each edge connects two different vertices and no two edges connect the same pair of vertices. This means a vertex cannot have a loop, i.e. an edge from a vertex to itself. \cite{bondy1976graph}

The neighborhood of a vertex, denoted $N_G(v)$, is a set of all vertices adjacent to $v$. A \emph{closed} neighborhood consists of all vertices adjacent to $v$ and $v$ itself. Unless stated otherwise, we will assume when talking about neighborhoods that they are closed neighborhoods. The degree of a vertex, denoted $d_G(v)$, is the number of vertices adjacent to $v$. The maximum degree of a graph $G$, denoted $\Delta(G)$, is the maximum degree of its vertices.

\subsection{Graph Coloring}
\label{sec:coloring}
A vertex coloring is an assignment of colors to each vertex of a graph $G$ such that no adjacent vertices share the same color. Mathematically, it can be described as a function $f : V \rightarrow S = \{1, 2, \dots, k\}$ such that $\forall u,w \in V$, if $(u,w) \in E$, then $f(u) \neq f(w)$. The \emph{chromatic number} of $G$, denoted $\chi(G)$, is the minimum number of colors, i.e. $|S|$, needed to properly color $G$.

\subsection{Complexity Theory}
\label{sec:complexitytheory}

>>>>

A graph, denoted $G=(V,E)$, is an ordered pair of two sets: a set of vertices $V$ and a set of edges $E$. Each edge consists of a pair of vertices from $V$. For example, $(u,w) \in E$ is an edge connecting vertices $u$ and $w$ where $u,w \in V$. Vertices are adjacent if they are connected by an edge. A graph can be non-technically described as a set of points (vertices) with lines (edges) connecting them. A simple graph is an undirected graph where each edge connects two different vertices and no two edges connect the same pair of vertices. This means a vertex cannot have a loop, i.e. an edge from a vertex to itself.

The neighborhood of a vertex, denoted $N_G(v)$, is a set of all vertices adjacent to $v$. A \emph{closed} neighborhood consists of all vertices adjacent to $v$ and $v$ itself. Unless stated otherwise, we will assume when talking about neighborhoods that they are closed neighborhoods. The degree of a vertex, denoted $d_G(v)$, is the number of vertices adjacent to $v$. The maximum degree of a graph $G$, denoted $\Delta(G)$, is the maximum degree of its vertices. \cite{bondy1976graph}

\subsection{Graph Coloring}
\label{sec:coloring}
A vertex coloring is an assignment of colors to each vertex of a graph $G$ such that no adjacent vertices share the same color. Mathematically, it can be described as a function $f : V \rightarrow S = \{1, 2, \dots, k\}$ such that $\forall u,w \in V$, if $(u,w) \in E$, then $f(u) \neq f(w)$. The \emph{chromatic number} of $G$, denoted $\chi(G)$, is the minimum number of colors, i.e. $|S|$, needed to properly color $G$. The \emph{vertex coloring} problem, when given a simple graph $G$, is to find $\chi(G)$. \cite{bondy1976graph}

A graph $G$ is said to be \emph{k-colorable} if it can be colored using $k$ or fewer colors, i.e. $\chi(G) \leq k$. A graph having $\chi(G) = k$ is said to be a \emph{k-chromatic} graph. The \emph{k-colorability} problem asks if a graph can be colored using $k$ colors. This problem is a slightly easier problem than the vertex coloring problem as it is decision problem rather than an optimization problem.

A \emph{conflict-free coloring} of a graph assigns one of $k$ colors to some of the vertices such that, for every vertex $v$, there is a color that is assigned to exactly one vertex among $v$ and $v$'s neighbors. A \emph{conflict-free k-coloring} of a simple graph $G$ assigns one of $k$ different colors to a subset $S \subseteq V$ of vertices such that $\forall v \in V$, there is a vertex $u \in N(v)$ where the color of $u$ is unique in the closed neighborhood of $v$. The vertex $u$ can be thought of the \emph{conflict-free neighbor} of $v$. The \emph{conflict-free chromatic number} of G, denoted $\chi_{CF}(G)$, is the smallest $k$ for which a conflict-free coloring exists. \cite{abel2017three}

\subsection{Computational Complexity Theory}
\label{sec:complexitytheory}
A \emph{decision} problem, as mentioned earlier, is a problem that can be answered with a `yes' or a `no' \cite{sipser2006introduction}. A decision problem is said to be in the class \emph{P} if in the worst case, it can be solved with an algorithm that runs in polynomial time. A problem can be solved in polynomial time if an algorithm with input size $n$ can run in at most $n^k$ steps where $k$ is a constant that does not depend on $n$.

Given a decision problem and information showing what the answer is, it can be possible to verify the answer quickly. If a decision problem can be verified in polynomial time but not solved in polynomial time, it is said to be in the class \emph{NP}. Take note that this does not exclude problems in class \emph{P}; \emph{P} is a subset of \emph{NP}. The \emph{k-colorability} problem is an example of a problem in NP. There are currently no known polynomial-time algorithms to solve the VCP but it can be verified in polynomial time. \cite{garey2002computers}

There are certain problems that can be proven to be as hard as every problem in NP. There problems are said to be NP-hard. A problem is NP-hard if it every problem in NP can be polynomially reduced to it. This brings an important result: if an NP-hard problem can be solved with an algorithm that runs in polynomial time, then any problem in NP could be solved in polynomial time. A decision problem is said to be \emph{NP-complete} when it is both in NP and NP-hard.

The vertex coloring problem (VCP) and the conflict-free coloring problem (CFCP) are not in the class NP as they are not decision problems. A common approach to proving problems as NP-hard or NP-complete is reducing them from a known NP problem. If we find a known hard problem $Y$, then we can prove that another problem $X$ is hard by reducing $Y$ to $X$. The VCP and the CFCP have both been shown to be NP-hard \cite{abel2017three,moret1998theory}. The \emph{k-colorability} problem is proven to be NP-complete with a reduction from 3-SAT, a well-known NP-complete problem \cite{sharma2012new}.
